### 2018-07-15 paper2/paper2.tex
onward and upward

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{}
\usepackage{tikz}

\usepackage[citestyle=authoryear,style=authoryear,backend=biber,natbib]{biblatex}
\addbibresource{paper2.bib}

\title{Military mechanization as a complex system}

\author{Tom Wallace}

\begin{document}

\maketitle

\begin{abstract}

	Military mechanization is the degree to which a state's army is comprised of
	ground combat vehicles as opposed to infantry. Mechanization is an important topic 
	in international relations scholarship and a practical concern for
	policymakers. This paper makes novel theoretical contributions by
	re-envisioning mechanization as a complex system. It also implements an
	agent-based model of military mechanization. The results of this model
	challenge the conventional wisdom that mechanization is primarily driven
	by external factors such as security environment. They imply that
	some internal factors (e.g. military culture or economic constraints)
	act as a ``brake'' on mechanization.

\end{abstract}

\newpage

\section{Introduction}

\subsection{Outline}

This paper is divided into five sections. This \textbf{Introduction} gives
background on military mechanization. The \textbf{Theory} section rigorously
develops mechanization as a complex system best represented by
agent-based modeling. The \textbf{Model} section presents a hypothesis and
implements an agent-based model in Python to test that hypothesis.
\textbf{Results} are shown. Implications are explored in \textbf{Discussion}.

\subsection{Background}

States differ in the composition of their ground forces.\footnote{States may have multiple organizations that
conduct ground combat operations: e.g., the United States operates both an Army
and a Marine Corps as independent services. Henceforth, this paper uses the
terms ``army'' and ``ground forces'' interchangeably.} One notable axis of comparison 
is \textit{mechanization}: the degree to which an army consists of infantry 
vs. combat vehicles. Treating mechanization as a continuous variable, 
one extreme is an army consisting exclusively of dismounted, small 
arms-wielding infantry units. The other extreme is an army consisting 
solely of tanks, armored personnel carriers, and the like.
No modern state fully embodies either extreme, but the 1970s-era Vietcong are
a good example of a low-mechanization army and the 1980s-era Israeli army of
high mechanization.

Mechanization is important for multiple academic literatures. A
fundamental debate in international relations (IR) is whether state
behavior is driven by structural insecurity and 
competition \citep{waltz2001man, mearsheimer2001tragedy}; or, whether other
factors such as domestic political structure \citep{russett1994grasping} and
socially constructed beliefs \citep{wendt1992anarchy} can point states to more peaceful directions. 
The extent to which patterns of mechanization do or do not correlate with
\textit{realpolitik} preparation for war informs this debate.
The security studies community is similarly split on the question of military
change. Some view militaries (particularly large ones) as change-averse, 
inertia-driven bureaucracies that can only be bestirred by major events (e.g.,
war or political reform)
\citep{posen1986sources,zegart2000flawed,locher2004victory,halperin2007bureaucratic}. 
Others contend that militaries can and do radically self-transform in both composition and doctrine
\citep{murray1998military,adamsky2010culture}. The patterns of
mechanization---stable? punctuated by sudden change? constantly evolving?---thus
have strong implications for this debate.

Mechanization choices also have more visceral impact.
\citet{lyall2009rage} find that highly mechanized militaries are less effective
at counterinsurgency, while \citet{biddle2004military} argues that
mechanization-enabled mobility aids conventional warfighting. It thus
pays to have an army that is designed to win the kinds of conflicts that a state is 
likely to face. Incorrect forecasting and choosing the wrong mechanization
posture costs blood and treasure. The experience of the U.S. military in the early
2000s, when it shifted under duress from a highly mechanized Cold War-legacy force to a less
mechanized counter-insurgency one, provides a vivid example. A quote from then-Secretary of Defense 
Donald Rumsfeld emphasizes the point:
``you go to war with the army you have, not the army you might wish to have'' \citep{schmitt_2004}.

There are many potential causes of mechanization. Inter-state
competition and war is a prominent candidate. States may choose choose mechanization policies 
that they believe will allow them to prevail against potential adversaries, or to send a
signal of deterrence, or to learn from the perceived mistakes of the last war
\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. 
Economic endowments may play a role: e.g., capital-rich states may mechanize more
than labor-rich states \citep{gartzke2001democracy}. Domestic political
institutions may matter for military composition: e.g., autocracies may be more willing than democracies to adopt
low-mechanization ``human wave'' tactics
\citep{quinlivan1999coup,reiter2002democracies,brooks2008shaping,talmadge2015dictator}. 
Ideology \citep{van1984cult} and national culture \citep[e.g.][]{pollack2004arabs} are
hard-to-measure but perhaps influential. Lastly,
organizational culture can play a role \citep{builder1989masks}. Militaries may pursue a
certain mechanization posture because it aligns with their institutional
objectives, which may or may not be coincident with rational strategy.

\citet{sechser2010army} is the landmark study on military mechanization. They
assemble a longitudinal dataset at the country-year level of the mechanization
of most states between 1979 and 2001. They conduct regression analysis on this
panel data. They use a ordinary least squares (OLS) linear regression model with corrected standard
errors,

\begin{equation}
y_{ij} = \beta_0 + \sum_{k=1}^P \beta_{ijk}x_{ijk} + \epsilon
\end{equation}

with $y_{ij}$ the mechanization level of state $i$ in year $j$ (measured as the 
ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
x_j \ldots x_{P}$ covariates hypothesized to be associated with
mechanization. They conclude that, out of the
aforementioned potential causes, ``choices about
mechanization are strongly associated with a state's security environment.''
Their dataset and conclusions have guided further scholarship on military
mechanization, e.g. \citet{pilster2012democracies}, which mostly follow the same
statistical approach.

\newpage

\section{Theory}

Definition of complex system. 
The orthodox multivariate statistical modeling approach is insufficient because
mechanization is a complex system. Agent-based modeling (ABM) is a more 
appropriate paradigm for studying it. These three
claims---insufficiency, complexity, and ABM---are addressed in turn.

\subsection{Inadequacy of statistical modeling}

\citet{sechser2010army} use observational rather than experimental
data. The assignment of treatments to subjects is not randomized, causing
\textit{selection bias}. For example, they operationalize
dyadic ``rivalry'' as whether two states recently engaged in a Militarized
Interstate Dispute (MID), and find significant association with mechanization. 
But interstate conflict is not a randomly assigned treatment. Rather, at least one state must select
itself into it \citep{fearon2002selection}. States differ in their
proclivity for conflict, based on many factors. 
Thus, ``rivalry'' is in fact
bundling together many uncontrolled covariates, casting doubt on the claimed
finding. Similar problems arise for other variables included in the model.
Natural experiments---a situation in which treatment is
known to be randomly assigned---are the typical econometric strategy to avoid selection bias. For
example, \citet{angrist2002vouchers} exploit the fact that private school
vouchers were distributed by lottery in Colombia to obtain unbiased estimates of
the effect of schooling type (private vs. public) on educational achievement.
But such situations are rare in IR: e.g., random assignment of war is an
absurd proposition. \citet{sechser2010army} lack a good control
because it probably does not exist. Reliance on selection-riddled
observational data means that econometrics are at best an incomplete
tool for studying mechanization.

\subsection{System complexity}

\subsubsection{Decomposibility}

The additive linear model in (1) assumes that each covariate $x$ is separable. Take as an example 
GDP per capita.\footnote{The example works as well with any
other covariate.} \citet{sechser2010army} find it negatively related with mechanization at the $p < 0.05$ level. 
But a valid interpretation requires an assumption
of its independence from other covariates (i.e. no multicollinearity).
It also makes an implicit claim about causal direction: that
mechanization $y$ is a function of various
$x$, and not the other way around.\footnote{To be clear,
\citet{sechser2010army} do not claim causation in their text. But
their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
direction of association.} Put differently, \textit{it requires that the
observed system can be neatly 
decomposed into constituent parts}. Figure 1 is a causal diagram depicting this logic. 
\begin{figure}[h!]
	\centering
	\caption{Decomposable system}
	\begin{tikzpicture}
		\node[draw, rectangle](Mechanization)  at  (0,0) {Mechanization};	
		\node[draw, rectangle](GDP per capita)  at  (4,0) {GDP per capita};
		\node[draw, rectangle](Other covariates)  at  (4,-2) {Other covariates};
		\draw[->] (GDP per capita) -- (Mechanization);
		\draw[->] (Other covariates) -- (Mechanization);
	\end{tikzpicture}
\end{figure}

Initial statistical testing seems to support this assumption. The model's variance
inflation factor (VIF) does not indicate multicollinearity.\footnote{See the Appendix for
a supporting table. Some coefficients have VIF $> 8$, which strongly
indicates multicollinearity. However, this only occurs with ``dummy'' variables
and the variable to which they relate, and so is not problematic.} However, VIF
is contingent on the specific model being tested, and evidence suggests that 
(1) is not a correctly specified model. It over-simplifies or omits key
relationships.

A large literature suggests that GDP per capita is associated with
regime type (e.g., democracy vs. autocracy), and causality appears to flow in
both directions. Democracy causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
Regime type affects the probability of war: the closest thing to an empirical law in international relations is
that democracies do not fight each other
\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. 
Regime type is widely argued to influence
battlefield performance in war because it affects the quality of military institutions and hence
capability of fielded forces \citep{reiter2002democracies,weeks2014dictators,talmadge2015dictator}. 
Battlefield performance affects regime type, e.g. a losing side may be subject to foreign-imposed
regime change a la 1939 France or 2003 Iraq
\citep{werner1996absolute,lo2008ensuring,downes2013forced}. \citet{sechser2010army} claim that the observed
historical probability of conflict (``rivalry'', in their terminology) affects mechanization; and claim that observed
battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects
battlefield performance in war; if it did not, we would not care about it
\citep[see also][]{lyall2009rage}.
Mechanization may affect GDP per capita in that states can over-spend on defense and 
impoverish the country
\citep{cappelen1984military,ward1992sizing,abu2003government}. Deterrence theory requires that 
battlefield performance affect the probability of war in that militarily 
potent states are unlikely to be attacked. The exact truth value of these claimed relationships is
uncertain---in no small part due to the limitations of econometrics that were just addressed---but
they are generally uncontroversial statements in their respective fields.

\begin{figure}[h]
	\centering
	\caption{Complex system}
	\begin{tikzpicture}
		\node[draw, rectangle](Mechanization)            at  (0,0)  {Mechanization};	
		\node[draw, rectangle](GDP per capita)           at  (4,0)  {GDP per capita};
		\node[draw, rectangle](Democracy)                at  (8,0)  {Democracy};
		\node[draw, rectangle](Probability of war)       at  (0,-4)  {Probability of war};	
		\node[draw, rectangle](Battlefield performance)  at  (8,-4)  {Battlefield performance};	
		\draw[->] (Mechanization) to (GDP per capita);
		\draw[->] (GDP per capita) to (Democracy);
		\draw[->] (Democracy) to (Battlefield performance);
		\draw[->] (Battlefield performance) to (Probability of war);
		\draw[->] (Probability of war) to (Mechanization);
		\draw[->] (GDP per capita) to[out=90,in=90] (Mechanization);
		\draw[->] (Democracy) to[out=270,in=270] (GDP per capita);
		\draw[->] (Battlefield performance) to[out=360,in=360] (Democracy);
		\draw[->] (Mechanization) to (Battlefield performance);
		\draw[->] (Mechanization) to[out=180,in=180] (Probability of war);
		\draw[->] (Democracy) to (Probability of war);
	\end{tikzpicture}
\end{figure}

Figure 2 is the logic diagram resulting from the above. It must be emphasized that we are only scratching the 
surface of system complexity. For example, Figure 2 does not address the nature of relationships 
between covariates: linear or non-linear? subject to multi-way
interactions? necessity and sufficiency distinctions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
not tractable to decomposition}. Ascertaining cause and effect in this system
through simple observation is impossible because of endogeneity. Moreover, controlling for
confounding variables breaks down when the controlled-for variable directly
relates to the outcome of interest \citep[``post-treatment bias'',
c.f.][]{king2010hard}. One would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
of the nodes and observe the resultant system state. But, as previously covered, one generally cannot conduct
experimental trials in international relations.  A modeling paradigm that would let us do so is needed.

\subsubsection{Cognition}

There is no law of nature dictating that an increase in one state's mechanization causes an adversary to 
raise its own mechanization. It also is not the case that many individual choices aggregate into something 
akin to natural law---the logic used 
in economics to justify treating supply and demand in a physics-esque manner---because the number of
people that decide mechanization choices is small. In a personalistic
dictatorship, it might be $n=1$. Even in polities with more dispersed authority, such decisions
typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
\textit{A state's mechanization level changes because someone decides to change it}.

If one takes the previous statement to be true, then a model of military
mechanization must be at least partly a cognitive model. Some agent or set of agents perceives the outside
world; processes that information; and takes action. Orthodox studies of
mechanization incorporate cognition to some degree. For example,
\citet{sechser2010army} test for learning: whether states that fail in
counter-insurgency wars decrease their mechanization as a result of this
``anti-mechanization lesson.'' This depiction, however, fails to capture 
key phenomena.

Human beings learn not only from their own experiences, but also from observing others.
State decisionmakers are no exception. For example, the stunning success of
Coalition precision-guided munitions in the 1991 Gulf War led other states who
were not direct combatants to reform their own militaries
\citep{scobell2011chinese}. There is no guarantee that such emulation always
works out. Consider Ireland: between the 1920s and 1940s, it sought to spend a
large portion of its defense budget on building a small number of
tanks, which at that time were novel technology being implemented by the leading
European militaries.
They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
but tanks were seen as a prestigious signifier of a ``real'' professional
military and hence desirable
\citep{farrell1998professionalization,farrell2001transnational}. Such learning
is absent from orthodox models of military mechanization.

\subsection{The utility of agent-based modeling}

In light of the above, agent-based modeling (ABM) is 
uniquely well-suited to study mechanization. Computer simulations provide the possibility 
to conduct controlled experiments \textit{in silico} \citep{epstein1996growing} that are not possible 
in the real world, freeing us from the tyranny of selection bias. Because the
modeler can adjust various ``dials'' and observe the outcome, ABM allows causal
inference or at least insight even in highly complex, perhaps unsolvable systems
\citep[c.f.][]{axtell2000agents}. ABM focuses on autonomous agents with
cognition, for which even relatively simple behavioral rules can generate
behavior at least as complex as mechanization
\citep{gilbert2005simulation,miller2009complex}.

Complex systems and ABM have been used in IR. Many methodology-focused
works have identified their utility to the field \citep[some prominent examples
are][]{modelski1990world,jervis1998system,pepinsky2005agents,de2014agent}.
Although actual usage has somewhat lagged behind theoretical enthusiasm,
ABM has been used to explore emergent 
balance of power \citep{cederman1997emergent}, international negotiation
\citep{earnest2008coordination}, and international crises and conflict
\citep{masad2016dissertation}. It is especially popular for studying
population-centric security phenomena such as terrorism and insurgency
\citep{cioffi2010mason,bhavnani2012modeling,weidmann2013violence,bhavnani2014group}.

The author's sense is that adoption of ABM in IR has lagged because it is
regarded as non-empirical. ABM offers no guarantees: just
because one's model ``works'' (internally consistent, externally plausible) 
does not mean it is true. Econometrics are the dominant methodological approach 
in quantitative IR because they seem to offer an objective yardstick for competing hypotheses: 
does A or B best fit the data? This paper contends that
epistemological limitations (most importantly, selection bias stemming from
non-randomization, and system complexity) mean that econometrics do not
provide reliable statistical inference for certain classes of IR problems, among
them mechanization \citep[c.f.][for a concurring opinion]{king2010hard}. 
If neither econometrics nor ABM can offer 
certitude that a hypothesis is true---as is the case for mechanization---then the latter 
should be preferred for its advantages in experimentation, transparency, researcher control, 
and amenability to complex systems.

\subsection{Research question and strategy}

I hypothesize that the arguments of \citet{sechser2010army} are incomplete and
that some omitted variable or set of variables acts as a ``brake'' on 
mechanization.\footnote{By ``omitted'', I mean that Sechser and Saunders conclude that the variable
is not very important for mechanization, not that they do not address the
variable at all.} An agent-based model is used to test this hypothesis. In broad
terms, agents (states) are defined to have the cognitive and learning behavior
implicitly claimed by \citet{sechser2010army}. The model is initialized with
real-world data from 1979 and allowed to run until 2001. The micro- and
macro-outcomes are then validated against real-world 2001 data. If
\citet{sechser2010army} are correct, the model's 2001 outcomes should match
real-world 2001 outcomes. If the model mechanization levels in 2001 are much
higher, then their claimed behavioral rules are incomplete and fail to include
some ``brake'' on mechanization, matching my hypothesis. Proving exactly what
that missing variable or set of variables is is outside the scope of this model
and paper, though it is discussed.

\section{Model}

This section of the paper describes the model. It does not use the ODD protocol
of \citet{grimm2006standard} but covers many of the same points.

\subsection{Agents}

There is one agent type: states. Agents exist at a fixed location and do not
move. The set of agents encompasses all diplomatically-recognized 
states in the world at a particular point in time, minus microstates (defined as
those with a population less than 750,000). Some other states were excluded; for
why, see the \textbf{Time}
section. Figure 1 depicts the set of agents.

\begin{figure}[h!]
	\centering
	\caption{Agents}
	\includegraphics[scale=0.05]{map.png}
\end{figure}

Agents have attributes. \textit{Name} is self-explanatory.
\textit{Mechanization} is the agent's level of mechanization at a given
timestep. Mechanization is calculated as the number of armored vehicles per 100
soldiers, per \citet{sechser2010army}. \textit{Neighbors} is a list containing
the \textit{Names} of an agent's geographic neighbors at a given point in
time---more on this in the subsequent \textbf{Environment} section. \textit{NeighborsMech} is the
average mechanization level of an agent's \textit{Neighbors} at a given point in
time. \textit{Enemies} is a list containing the \textit{Names} of an agent's
enemies at a given point in time. An enemy is an agent with which the agent in
question has had a militarized interstate dispute in the preceding ten years,
per the Correlates of War (COW) Militarized Interstate Disputes Database (MIDB)
\citep{cow_midb}. \textit{EnemiesMech} is the average mechanization score of
enemies. Agents also have attributes that affect their
perception of other agents and cognition for changing their own behavior. These
attributes are discussed in the \textbf{Cognition and Actions} section.

\subsection{Environment}

Agents exist in a spatial environment. Spatiality is represented by neighbor
relationships. Two states are defined as
being neighbors if they share a contiguous land border or a stretch of water
less than 400 miles. Data on neighbor relationships is taken from the Correlates
of War (COW) Direct Continuity dataset \citep{cow_contiguity}.
Modeling spatiality as dyadic neighbor relationships, rather than a more
detailed raster or vector GIS implementation, is justified on the basis that the
academic literature on mechanization suggests that two states neighboring each
other is the important factor, not more detailed geographic information (e.g.,
topographical features along the border). As such, this paper's modeling choice
captures the phenomenon of interest while avoiding the overhead of a full-blown
GIS implementation coupled with ABM \citep[e.g.][]{harper2002modeling,bennett2006modelling}.

\subsection{Cognition}

Agents have global perception of other agents: they can ``see'' the
mechanization state of all other agents. This modeling choice is reflective of
the real world, in which states devote significant resources to stay informed of
the military posture of others. Agent cognition centers on 
updating their own mechanization in response to the external security
environment. This occurs in three ways. One, if an agent's
\textit{NeighborsMech} increases---i.e., the agent's neighbors become more
mechanized---the agent will increase its own mechanization by some proportional
amount controlled by the agent's unique \textit{NeighborMechSensitivity}
attribute, which is heterogenous across agents (more on this below).
Two, if an agent's
\textit{EnemiesMech} increases---i.e., the agent's enemies become more
mechanized---the agent will increase its own mechanization by some proportional
amount controlled by the agent's unique \textit{EnemyMechSensitivity} attribute,
again heterogenous across agents. Three,
agents occasionally suffer an ``anti-mechanization lesson'': defeat in a
counter-insurgency conflict. Such events are scripted to occur for agents and
years as per the historical record: e.g., the Soviet Union experienced such a
lesson in the 1980s in Afghanistan. An agent experiencing such an event will
decrease their own mechanization by some amount controlled by the agent's
unique \textit{LessonSensitivity} attribute, again heterogenous across agents.
Cognition for agent $i$ thus is:

\begin{multline}
	\textrm{Mech}_i^{(t+1)} = \textrm{Mech}_i^{(t)} \\ 
	+ (\textrm{NeighborMechSensitivity}_i \times \textrm{NeighborMech}_i^{(t)}) \\
	+ (\textrm{EnemyMechSensitivity}_i \times \textrm{EnemyMech}_i^{(t)}) \\
	+ (\textrm{LessonSensitivity}_i \times I_{\textrm{Lesson}}(t, i))
\end{multline}

with $(t)$ denoting timestep, $i$ indexing agents, and $I_{\textrm{Lesson}}(t,
i)$ an indicator function of whether the agent experienced an anti-mechanization
lesson. Table 1 summarizes agent attributes.

\begin{table}[h]
	\centering
	\caption{Attributes, organized by theme}
	\begin{tabular}{|l l|}
		\hline
		\textbf{Identity} & \textit{Name} \\
		& \\
		\textbf{Mechanization} & \textit{Mechanization} \\
		& \\
		\textbf{Spatiality} & \textit{Neighbors} \\
		& \\
		\textbf{Inter-agent Relationships} & \textit{Enemies} \\
		& \\
		\textbf{Perception} & \textit{EnemiesMech} \\
		& \textit{NeighborsMech} \\
		& \textit{Lesson} \\
		& \\
		\textbf{Cognition} & \textit{EnemyMechSensitivity} \\
		& \textit{NeighborsMechSensitivity} \\
		& \textit{LessonSensitivity} \\
		\hline
	\end{tabular}
\end{table}

An important note is that these three factors---neighbors, enemies, and
defeat in counter-insurgency---correspond to the factors identified by 
\citet{sechser2010army} as most important for mechanization. Because the research
goal is essentially to see whether their findings stand up to agentization,
it is critical to make agent cognition work in about the way that their findings
predict. Inclusion of other factors not identified as important for
mechanization---e.g., alliances, economic factors,
culture, etc.---would be interesting but counter-productive to the research
question. They are discussed along with this model's findings.

Data for \textit{NeighborsMechSensitivity}, \textit{EnemiesMechSensitivity}, and
\textit{LessonSensitivity} come from \citet{sechser2010army}. Specifically, they
estimate coefficients and standard errors for these three factors, thus defining
a statistical distribution for each. Upon instantiation, every agent makes a
random draw from each of the three distributions, the resulting values of are
taken for these three attributes. This modeling choice has two important
elements. One, it further enhances agent heterogeneity. Agents already differ in
their geographic neighbors, enemies, and experiences with anti-mechanization
lessons; now, they also differ in the cognition used to update mechanization
based on these factors. It also introduces a stochastic element to the
model, meaning that each run differs and so multiple runs must be conducted for
each specification.

\subsection{Time}

The model starts in 1979 and runs in 2-year time increments until 2001. This
timespan and periodicity reflects data availability: a longer time period and
more frequent updates would be ideal but is not practically achievable. Agent
attributes are initialized, and then then the model iterates through timesteps. 
In discussing what happens each timestep, it is useful to distinguish between
things that depend on agents' actions and those that do not.

Some changes are deterministically programmed into the model and do not depend on
agent behavior. Every timestep, neighbor relationships and enemy relationships
are updated in accordance with the historical record. The occurrence of
counter-insurgency defeats also is hardwired into the model and does not depend
on agent behavior within the model.

Other changes do depend on agent behavior. Agent activation order in a
particular timestep is random. Every timestep, every agent perceives the world
aorund them; recalculates \textit{NeighborsMech} and \textit{EnemiesMech};
updates their mechanization level according to \textit{NeighborMechSensitivity}
and \textit{EnemyMechSensitivity}; and updates their mechanization level
according to \textit{LessonSensitivity} if an anti-mechanization lesson occured.

An important caveat is that the set of agents represented in the model does not
change over time: i.e., agents are the same from 1979 onwards. In real life,
of course, the agent population was not static over this time period, with many
states ceasing to exist in 1991 and new ones forming. Modeling these changes is
beyond the scope of this model. How a collapsed state's military assets are divvied up
among successor states is complex and not tractable to simple representation
(e.g. averaging). For example, the Soviet Union deployed most of its modern
military equipment in the far west of the country in anticipation of a conflict
with NATO, and so when the USSR collapsed Ukraine inherited the lion's share, leaving
little for successor states such as Azerbaijan that were not important military
districts \citep{roy1993military}. Such inheritance has
a strong effect on the mechanization of of successor states. Designing
complex and state-specific ``rules of inheritance'' was too much for this
initial effort but is an area for further development.

\subsection{Output Data, Verification, and Validation}

The main model output of interest is data on military mechanization. It is provided for
every agent and every time-step. These agent-level data also can be aggregated
into system-level measures of military mechanization. Per the research design,
the intent is to compare the model output to what actually occurred. If the
model produces military mechanization levels much higher than reality, it means
that Sechser and Saunder's findings do not stand up to agentization and there is
some missing ``brake'' on military mechanization. Note that this research
strategy has validation (comparison of model output to empirical data) at its
core and so no separate validation is needed. Verification that the model is
doing what it is intended was accomplished through manual inspection.

\section{Results}

At both the system- and agent-level, this agent-based model results in much
higher mechanization levels than occurred in the real world. First, let us
compare the average mechanization levels across all countries at each time-step
of the model. Although the set of agents is not exactly the same between the
model and reality (c.f. earlier issues with inheritance), this
gross comparison still is a useful measure of comparison. Model results are
the average of 100 runs, with all parameters (e.g.
\textit{EnemyMechSensitivity}) drawn from the distributions claimed by
\citet{sechser2010army}; empirical results are drawn from the National
Mechanization Index compiled by \citet{sechser2010army}.

\begin{table}[h]
	\centering
	\caption{Comparison}
	\begin{tabular}{|l r r r|}
		\hline
		\textbf{Year} & \textbf{Model} & \textbf{Reality} & \textbf{Difference}\\
		1979 & 0.0156 & 0.0139 & 0.0017 \\
		1981 & 0.0189 & 0.0167 & 0.0022 \\
		1983 & 0.0229 & 0.0167 & 0.0062 \\
		1985 & 0.0280 & 0.0183 & 0.0097 \\
		1987 & 0.0351 & 0.0187 & 0.0164 \\
		1989 & 0.0415 & 0.0197 & 0.0218 \\
		1991 & 0.0512 & 0.0215 & 0.0297 \\
		1993 & 0.0636 & 0.0229 & 0.0407 \\
		1995 & 0.0804 & 0.0231 & 0.0573 \\
		1997 & 0.1014 & 0.0270 & 0.0744 \\
		1999 & 0.1319 & 0.0272 & 0.1047 \\
		2001 & 0.1730 & 0.0291 & 0.1439 \\
		\hline
	\end{tabular}
\end{table}

\newpage

Examining results at the agent-level reveals that the higher mechanization
levels of the model are much more pronounced for some agents than others. In
some cases, the model actually results in \textit{lower} mechanization levels,
with Mexico serving as one example.

\begin{table}[h]
	\centering
	\caption{Mexico}
	\begin{tabular}{|l r r r|}
		\hline
		\textbf{Year} & \textbf{Model} & \textbf{Reality} & \textbf{Difference}\\
		1979 & 0.0013 & 0.0012 & 0.0001  \\
		1981 & 0.0014 & 0.0012 & 0.0002  \\
		1983 & 0.0018 & 0.0026 & -0.0008 \\
		1985 & 0.0019 & 0.0013 & 0.0006  \\
		1987 & 0.0021 & 0.0019 & 0.0002  \\
		1989 & 0.0022 & 0.0027 & -0.0005 \\
		1991 & 0.0024 & 0.0030 & -0.0006 \\
		1993 & 0.0026 & 0.0033 & -0.0007 \\
		1995 & 0.0030 & 0.0035 & -0.0005 \\
		1997 & 0.0032 & 0.0074 & -0.0042 \\
		1999 & 0.0036 & 0.0082 & -0.0046 \\
		2001 & 0.0040 & 0.0075 & -0.0035 \\
		\hline
	\end{tabular}
\end{table}

For some agents, however, mechanization levels are extremely high. 
Mechanization score is a percentage, and so values over 1 do not make sense, 
yet some countries end the simulation with mechanization scores over 1. Note
that one could easily right-bound values at 1, but this was not done in order to
see what values would be organically converged to.

\begin{table}[h]
	\centering
	\caption{Highest mechanization scores, 2001}
	\begin{tabular}{|l r|}
		\hline
		\textbf{Agent} & \textbf{Mechanization Score} \\
		Israel & 1.59 \\
		USA & 1.34 \\
		Kuwait & 1.14 \\
		Russia & 1.06 \\
		\hline
	\end{tabular}
\end{table}

There appears to be a correlation between number of enemies and high
mechanization score. For example, both the USA and Russia consistently had some
of the highest amount of enemies, peaking with 16 enemies for Russia in 1999
(this high number is attributable to the Kosovo crisis, when a confrontation between Russian and NATO forces
at Pristina Airport nearly led to war), and as per the above table are among the
most mechanized states. That Israel had many enemies during this time period is
obvious.

There also appears to be neighborhood effects. Mechanization is not evenly
distributed across the globe. Rather, it is geographically concentrated in a few
hot-spots. Agents belonging to the Middle East and North Africa (MENA) region
account for about 10\% of the overall agent population, yet make up 50\% of the
top 20 most mechanized agents at simulation end. Other regions
are correspondingly under-represented; for example, Asia does not have a single
state in the top 20.

\begin{table}[h]
	\centering
	\caption{Top 20 mechanized agents at simulation end, by region}
	\begin{tabular}{|l l l l l l|}
		\hline
		\textbf{MENA} & \textbf{Europe \& FSU} & \textbf{Africa} & \textbf{Americas} & \textbf{Asia} & \textbf{Oceania} \\
		Israel       & Russia         & Sudan        & USA    & (none) & (none) \\
		Kuwait       & France         & S. Africa    & Canada &        &        \\
		Iran         & UK             &              &        &        &        \\
		Libya        & Greece         &              &        &        &        \\
		Iraq         & Netherlands    &              &        &        &        \\
		Syria        & Poland         &              &        &        &        \\
		KSA          &                &              &        &        &        \\
		Egypt        &                &              &        &        &        \\
		Lebanon      &                &              &        &        &        \\
		Turkey       &                &              &        &        &        \\
		\hline
	\end{tabular}
\end{table}

\section{Discussion}

\subsection{Findings}

Two conclusions should be drawn. One, the model results in much higher
system-level mechanization than occurred in reality. Two, this higher
mechanization is not evenly distributed across agents; rather, a small number of
agents with many enemies and that all border each other become outliers and
hence drive the global average far upward. The explanation for both phenomena
lies in a positive feedback loop. Enemies and neighbors adjust their
mechanization scores in response to each other; hence, a geographic cluster of
enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
mechanization.

Three, because this model is a
faithful representation of the factors that \citet{sechser2010army} claim are
most important for mechanization, that model output fails to match reality
indicates that their claimed model is incomplete. There must be some ``brake''
on mechanization; put differently, some countervailing factor(s) must mitigate the
effect of the positive feedback loop identified above. This model cannot
conclusively identify what that ``brake'' is. The most plausible candidates are
factors examined by \citet{sechser2010army} and deemed statistically
insignificant such as regime type and economy, and factors not examined by
\citet{sechser2010army} at all such as culture.

This paper is a novel contribution to the literature. Although agent-based
models have found traction in the study of grievance, insurgency, and
other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
they quite rarely are used to model international relations and state-centric security
phenomena despite some early work on this front
\citep[see][]{,axelrod1997complexity}. 
This paper's agentization of an orthodox security
studies model thus is novel in method. Its findings also are important. They
show that they show a solely ``external'', security
environment-focused view of mechanization does not stand up to agentization.
Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
loops---all hallmarks of a complex system---drive results that significantly
diverge from theoretical expectations of the orthodox view.

\subsection{Areas for Improvement}

Having taken a critical view of \citet{sechser2010army}, it only is fair that
this paper applies that lens to itself. There are two main areas for
improvement.

One, this model's representation of \citet{sechser2010army} could be tightened
up, as there currently is divergence between the two. For example, this model
examines a smaller agent set than their paper. Building ``inheritance''
processes to allow new states to emerge is desirable. This paper's
use of statistical coefficients from \citet{sechser2010army} also has some
methodological problems, a full description of which is omitted here due to
space constraints. It suffices to say that a better approach would be to conduct
more extensive parameter sweeps of relevant cognitive factors.

Two, this model could incorporate potential explanations of what the missing
``brake'' on mechanization might be. Economy and culture could be included in
the model. Doing so would allow exploration of whether these factors are
logically consistent and can act as a countervailing factor that controls the
positive feedback loop of mechanization. It cannot \textit{prove} that these or
any other factors actually exist in reality, but they can demonstrate how they
are or are not consistent in a simulated reality.

My hope is that making these changes will make this a publication-quality paper
capable of serving as one of two required papers for the comprehensive exams.

\newpage

\printbibliography[heading=bibnumbered]

In summary, the orthodox view of mechanization is that it is a
reflexive process: states choose their mechanization posture largely based on
that neighbors, enemies, and recent military experiences (e.g. defeat in
counter-insurgency). This paper argues
that this view---based on traditional longitudinal data analysis---is ripe for
agentization. It implicitly argues for autonomous agents (states) that have
cognition regarding mechanization that takes into account other agents (e.g.
enemies), spatiality (e.g. neighbors), and learning from experience. This paper
builds such a model and validates it against the empirical record, using the
National Mechanization Index of \citet{sechser2010army}. It finds that the 
orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
causes states to
mechanize much faster and to a greater degree than occurs in the real-world.
This implies the existence of some countervailing factor (possibly a negative
feedback loop) that mitigates the extent of mechanization. Candidates for what
this could be mostly center on ``internal'' factors---e.g., culture, economy,
and so on---as opposed to the ``external'', security environment-centric
orthodox view.

are ripe for agentization. The
conclusions drawn implicitly claim
that states are autonomous agents; that they have
cognition regarding how to perceive the outside world
and how to react to it; that spatiality and
neighborhood matter; that inter-agent relationships matter; and that
agents learn from experience. Agent-based modeling is uniquely well-suited to
represent such phenomena .

Importantly, the findings of \citet{sechser2010army} imply a positive feedback
loop. State A has high
mechanization, causing enemy State B to raise their mechanization, causing State
A to raise their mechanization... and so on. Such feedback loops are a hallmark
of complex systems and again are well-suited for representation by agent-based
modeling. This specific feedback loop is often called
the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.

\begin{table}[h]
	\centering
	\begin{tabular}{l r}
		\hline
		\textbf{Variable} & \textbf{VIF} \\ 
		\hline
		log\_mech\_enemies & 8.76 \\ 
		enemies\_dummy & 9.12 \\ 
		log\_mech\_neighbors & 2.19 \\ 
		neighbors\_dummy & 2.03 \\ 
		log\_mountainous & 1.26 \\ 
		antimech\_lesson\_2 & 1.21 \\ 
		promech\_lesson\_2 & 1.11 \\ 
		democracy & 1.69 \\ 
		military\_govt & 1.13 \\ 
		instability & 1.12 \\ 
		log\_gdp\_cap & 2.10 \\ 
		iron\_steel & 1.32 \\ 
		log\_mech\_allies & 12.64 \\ 
		allies\_dummy & 12.59 \\ 
		cold\_war\_bloc & 1.48 \\ 
		igos & 2.03 \\ 
		post\_soviet & 1.33 \\ 
		log\_mech\_lag2 & 1.57 \\ 
		\hline
	\end{tabular}
\end{table}
\end{document}